---
title: "Math 158 Project: Teach for America Data Assignment 2"
author: "Sophia Hui and Allison Kirkegaard"
date: "February 10, 2018"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, fig.align = "center")
require(dplyr)
require(ggplot2)
require(infer)
require(skimr)
require(broom)
require(haven)
require(readr)
TFAdata <- read_csv("my_TFAdata.csv")
options(digits=3)
```
##### Introduction

Our data comes from an evaluation of Teach for America conducted by Mathematica Policy Research in 2004, which was designed as a year-long randomized controlled trial including 17 elementary schools from six regions across the United States. It comprises student pre- and posttest scores in math and reading, student survey responses, and teacher survey responses. We chose to have students be our observational units, and merged their teachers' data with their own so that our model of student test performance can include both student and teacher characteristics.

##### Hypothesis Testing

We are introducing a new variable: the difference between pretest and posttest scores in math; the variable will be called 'diff'. We will be testing the linear relationship between 'diff' and the teacher's years of experience. Our explanatory variable will be total hours the teacher spent on various professional development (e.g., in-depth study of content, methods of teaching, applications of technology), and our response variable will be $X$ = posttest score - pretest score in math. The linear regression model we will be using is $$ Y_i = \beta_0 + \beta_1 X_i + \epsilon_i$$. Our hypothesis is:

$$ H_0: \beta_1 = 0$$
$$ H_a: \beta_1\neq 0$$

##### Checking Assumptions for Linear Regression

```{r}
TFAdata <- TFAdata %>% mutate(diff = ss_m-press_m)

ggplot(TFAdata, aes(x=a1_a, y=diff)) +
  geom_point() + 
  geom_smooth(method="lm", se=FALSE)

test_lm <- lm(diff ~ a1_a, data=TFAdata)
tidy(test_lm)
test_res <- resid(test_lm)
plot(fitted(test_lm), test_res, ylab="Residuals", xlab="Total Years of Teaching Experience", main="Teacher Experience and Difference in Math Test Scores Residual Plot")
abline(0,0)
```

Note that 100 observations were removed, because teachers did not provide how many years of experience they had in teaching, including the current school year. 

Because our explanatory variable is discrete, it is difficult to see a linear relationship between our X and Y. However, there is no curvilinear relationship present either, so we can assume that the relationship is linear with constant variability. There is independence among the residuals, and they are normally distributed. However, there is no constant variance, because at high number of years of experience, the residuals are larger. 

Our linear model is $$\hat{Y}=12.235 + -0.193X$$, with a p-value of 8.20e-04. Thus, we can reject $H_0$. 

##### Confidence Interval
```{r}
newtest <- data.frame(a1_a=1)
crit_val <- qt(.975, glance(test_lm)$df.resid)
test_pred <- augment(test_lm, newdata=newtest, type.predict = "response")
.se.pred <- sqrt(glance(test_lm)$sigma^2 + test_pred$.se.fit)
test_pred <- test_pred %>%
mutate(lower_PI = .fitted - crit_val * .se.pred,
upper_PI = .fitted + crit_val * .se.pred,
lower_CI = .fitted - crit_val * .se.fit,
upper_CI = .fitted + crit_val * .se.fit)
test_pred
```
By constructing a 95% prediction interval, we are 95% confident that the difference in pre- and post-test math scores will be in the interval (-15.6, 39.7) for a student who has a teacher with no prior experience ($X_h$ = 1). We also constructed a 95% confidence interval, which tells us that if we were to repeatedly samples, then 95% of our confidence intervals will capture the true mean score difference for students with teachers with 1 year of experience (including the current school year). 

##### Fit of Our Model
A concern about our residuals plot is the non-constant variance. When we look at the residuals around 12 years of teaching experience, their magnitudes are larger than the other residuals in the plot. 
```{r}
summary(test_lm)
```
Additionally, our $R^2$ value is very low. Only 0.924% of the variability of the response data is explained by our model. Therefore, a linear regression may not be the best model for this data. 

##### Conclusion
Initially, we were concerned by the scatterplot, because our explanatory variable only had discrete values, which yielded in a plot that did not look linear. However, we were able to reject our null hypothesis and conclude that there was a negative relationship between the total number of years of teaching experience and the difference in math pre- and post-test scores. This surprised us, because typically, more experienced teachers have better success in the classroom. Yet, because we were unable to meet the technical condition of having constant variance and because of our low $R^2$ value, we cannot confidently conclude anything about the relationship. 

